Model Recognition as Planning

### 2019-03-10 doc/icaps-19/camera-ready/model-recognition.tex
last update in conclusions

This paper formalizes the task of {\em model recognition} and introduces a novel method that estimates the probability of a \strips\ model to explain a partial observation of a plan execution. Our {\em model recognition as planning} approach builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and incomplete or missing intermediate states in the observation. Once the planning model of the observed agent is recognized, the model-based machinery for automated planning becomes fully applicable for other recognition tasks like {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} or {\em counter-planning}~\cite{PozancoEFB18}, which require beforehand a model of the observed agent.

We show the effectiveness of our approach in three experiments that encode a set of different \strips\ models.
The first experiment is a proof-of-concept \emph{classification task} that perfectly recognizes the regular automata of a given string. The second experiment, where the number of trajectories explaining the observation is bounded, shows that a more accurate model prediction is achieved with more complete observations. Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable. A significant lesson from the experiments is that our approach allows us to predict behaviours that are not ruled by a goal-driven deliberative planning model (e.g., models of fixed programs or finite state machines --regular automata-).

>>>>

This paper formalizes the task of {\em model recognition} and introduces a novel method that estimates the probability of a \strips\ model to explain a partial observation of a plan execution. Our {\em model recognition as planning} approach builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and incomplete or missing intermediate states in the observation. Once the planning model of the observed agent is recognized, the model-based machinery for automated planning becomes fully applicable for other recognition tasks like {\em goal recognition}~\cite{RamirezG10,ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} or {\em counter-planning}~\cite{PozancoEFB18}, which require beforehand a model of the observed agent.

We show the effectiveness of our approach in three experiments that encode a set of different \strips\ models.
The first experiment is a proof-of-concept \emph{classification task} that perfectly recognizes the regular automata of a given string. The second experiment, where the number of trajectories explaining the observation is bounded, shows that a more accurate model prediction is achieved with more complete observations. Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.

A significant lesson from the experiments is that our approach allows us to predict behaviours that are not ruled by a goal-driven deliberative planning model such as, for instance, models of fixed programs or finite state machines (regular automata).
